Efficiency

### 2018-09-26 Second.Rnw
Final Draft

However, over a number of years now the pumpjet choice has attracted sustained attention and criticism, mostly around its efficiency at low speeds, where a diesel-electric submarine would necessarily spend a great deal of its operating time.  I was commissioned by Gary Johnston of Submarines for Australia to investigate this topic in detail and produce a report \parencite{morrison2018}, which concluded that a very substantial reduction in dived range and endurance (as well as overall range) would necessarily result from the pumpjet choice.  Furthermore, that paper challenged the assumption that the pumpjet would be acoustically superior at low speeds, where open propellers would have no cavitation, and cause much less turbulence than the equivalent jet.

That paper, and a subsequent blog post \parencite{morrison2018a} on the topic which made the paper publicly available, attracted the interest of the Senate Economics Reference Committee Inquiry into the Future of Naval Shipbuilding in Australia.  Senator Chris Ketter asked written questions on notice to the Department of Defence requesting responses to the paper specifically, and related comments from public officials, with a return date of 3rd April 2018.  I was also invited to provide evidence to the committee in a public hearing on the 7th of June 2018.  In that public hearing \ref{SENmorrison}, I stood by the conclusion of my substantive paper, which was: ``between two otherwise identical submarines, the one with the pump jet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurance and a lower overall range than the one with the propeller."  In the presence of both the outgoing acting CEO of Naval Group Australia Brent Clark, and General Manager of Submarines (for Defence) Stephen Johnson, I pointed out that that statement was yet to be challenged, and that I'd be very surprised if anyone with technical competence or professional integrity sought to challenge it.

In their own evidence during the course of the day, neither of those two did choose to challenge it.  Senator Kim Carr requested that defence produce a written response to that report for the committee prior to their handing down of their final report for the Inquiry, and Defence agreed to do their best in the time available, with the head of Capability and Sustainment Group (CASG) Kim Gillis assuring the committee that ``We  will  provide  the  committee  with  the  best  possible  advice  we can  in  the  shortest  possible  time  frame  to  meet  your  requirements  so  that  you  can  be  better informed, because we do hold that expertise." Senator Carr emphasised the importance of timeliness that ``it would look pretty bloody stupid after the report's handed in''. Despite these exhortations, in the final report of the committee this request, and Defence's acceptance of it was noted, but no acknowledgement was made of whether any response was actually received, or what it contained \parencite{SERC2018final}.

In the August edition of Australian Defence Magazine a feature article on the titled `Future Submarine: Pumpjet propulsor technology choice still not well explained' \parencite{kerr2018} appeared, referencing both my testimony in the hearing and the paper.  Comment for the article was sought from CASG, the Royal Australian Navy (RAN) and Naval Group, and none was forthcoming.  However, some substantial comments were provided by Dr David Kershaw, the chief of DSTG's Maritime Division, who was also Science and Technology Advisor for a time on SEA 1000, the Future Submarine Project.

@@

In early September, the Department of Defence gave a six-page written response to my paper, a full three months after the initial request. The two substantive arguments already identified were repeated, but the weight of their argument actually came to rest on another assumption.  This assumption was mentioned already by Dr Kershaw in the ADM article, but it was so completely at odds with the most basic physics that I was inclined to dismiss it as perhaps a clumsily worded statement made in an interview compounded by a little journalistic mis-quotation, rather than a substantive argument.   The re-appearance of this argument, which was spelled out and repeated multiple times in a written document that was prepared by Defence over the course of three months, well after any nominal deadline for timeliness had elapsed, dispelled all illusions that the argument was included by accident.  As such, I have also addressed it here as the first of three substantial arguments made against my earlier conclusions regarding pumpjet efficiency, which has quite significantly altered the conclusions of this paper, which seeks to consider all three lines of reasoning in sufficient detail to suitably discuss the circumstances in which such arguments might be made.

@@

\section{A brief chronology of comment}

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{28 May 2012 - Senate Estimates}]
\textit{Senator Johnston directly challenges Rear Admiral Justin Moffit (then director of SEA 1000) about the pumpjet proposal.  Full exchange in \ref{SENMoffit}. Emphasis added.}
\\
\\
\textbf{Senator JOHNSTON:} I am told that there are no pump jets on any conventional class submarines in the world.
\\
\textbf{Rear Adm. Moffitt:} That is quite right. It might be a very easy task for us to write it off. I do not know, because I have not done-
\\
\textbf{Senator JOHNSTON:} The French and the Russians have tried. No-one has used them. They are not being used on anything other than nuclear powered submarines-particularly the Americans.
\\
\textbf{Rear Adm. Moffitt:} \textbf{If it does not look like it makes sense, I will rule it out. But I cannot give you the answer to that yet because we have not done the work.}
\\
\textbf{Senator JOHNSTON:} Who had the idea to look at pump jets? Where did that come from? Did it come from capability group saying that we had better had a look at that?
\\
\textbf{Rear Adm. Moffitt:} I do not know the answer to that. I doubt that there is a clear answer.

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{10 Feb 2016 - Sean Costello, CEO DCNS Australia}]
\textit{In a blog post\parencite{costello2016} for the ASPI Strategist prior to the announcement in April, Sean Costello describes the significance of the pump jet technology to their bid. Emphasis added.}
\\
\\
``As a complete submarine power, we understand conventional propulsion, which is why we also understand that propulsion is but one part of the submarine puzzle. In designing to the Australian requirement, it should come as no surprise then that \textbf{the conventionally powered Shortfin Barracuda Block 1A is only 5\% lighter than its nuclear cousin.}"
\\
\\
``There's no better example of what I'm talking about than the \textbf{offer from France to transfer to Australia sovereign control and use of pump-jet propulsion technology for the Shortfin Barracuda}-technology resident only in France, the UK and the USA. Technology born from the French SSBN program a generation ago. The stealth and hydrodynamic performances of pump jet propulsion are of course classified and in Australia known only to DCNS and the Australian Government."

@@

The significance of the pumpjet in Frances bid was clearly acknowledged well before the public announcement of DCNS's success.  It also provides a clear indication that the size of the submarine already appeared to be quite well-known, and was only intended to be a very minor modification from the nuclear Barracuda.  For such a small variation in displacement, it might be very reasonably expected that the length of the submarine would vary (the easiest modification), but the diameter would not need to.  These comments also display quite clearly that the pumpjet technology in France's possession is recognised as common with technology in the US and the UK.  This rules out the likelihood that some special or unique pumpjet technology, better suited just to diesel-electric submarines, had been pioneered by France alone.  This comment also directly implies that information regarding the performance of the pumpjet had already been given to the Australian  government, \textit{prior} to the conclusion of the Competitive Evaluation Process (CEP).

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{8 Apr 2016 - Sean Costello, CEO DCNS Australia, and Gerard Autret, Chief Naval Architect of the Shortfin Barracuda}]
\textit{In a blog post\parencite{autret2016} for the ASPI Strategist prior to the announcement in April, Sean Costello describes a number of the major features and design approaches for the Shortfin Barracuda, including relating the choice of the pumpjet closely to an acoustic requirement to close with another submarine at higher speed.}
\\
\\
``The challenge for any attack submarine is to maintain a nearly silent acoustic signature at a speed necessary to manoeuvre within weapons range of the target. The nuclear Barracuda is designed to \textbf{reduce radiated noise when operating at a speed sufficient in order to close a threat submarine undetected. Of particular importance is the pump-jet propulsor}, which combines a rotor and stator within a duct to significantly reduce the level of radiated noise through the effects of wake harmonisation and avoidance of cavitation."
\end{tcolorbox}

This comment provides clear context for the importance of the pump-jet.  Consistent with the physics outlined in Reference \parencite{morrison2018} the acoustic advantage is only emphasised by DCNS as being significant at higher speeds.  Whether it is viable for a diesel-electric submarine (as opposed to the nuclear one mentioned here) to maintain such speeds to close with many types of targets is left unclear, or at least unmentioned.  As later discussed in my evidence to the Senate \ref{SENmorrison}, this is consistent with DCNS being well aware with the advantages and disadvantages of the pumpjet, whilst not making verifiably false claims.

@@

The official explanation of the designer and builder's website makes a carefully worded explanation of the superiority of the pumpjet, by limiting the comparison to a ``tactical" encounter with an ``otherwise identical" submarine.  As discussed in \ref{SENmorrison}, this is the only \textit{possible} target that an elevated speed - such as could be achieved by a diesel-electric submarine - might assist in closing with.  The submarine would be at a substantial `strategic' disadvantage in such an encounter, since its lower dived endurance, dived range, and poorer indiscretion ratio dramatically increases the probability that the pumpjet propelled submarine would wind up being the hunted, rather than the hunter.  More significantly, the speeds achieved would not allow the submarine to close with any other target from behind such as almost all surface ships, or nuclear submarines.  This statement exhibits extreme care in the wording in order to be plausibly defensible, and is arguably misleading.

@@

@@

@@

\subsection{Naval Group at PACIFIC 2017 and in ASPI Strategist}
\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{4 October 2017 - Jean-Michel Billig, Executive Director Australian submarine Program, Naval Group}]
\textit{In comments to journalists at the PACIFIC 2017 trade show, the Naval Group executive in charge of the Australian Program acknowledges that conventional propellers may be used on the Shortfin Barracuda \parencite{sbs2017}.  Subsequent reports quote Naval Group shifting responsibility for the propulsion choice to the Australian Government.}
\\
\\
(Interview transcribed from Reference \parencite{lane2017}.)

@@

\textbf{Jean-Michel Billig:} Yes, yes, we could, we could, we could also end up with something different
\\
\\
``However decisions on forms of propulsion are ultimately a matter for the Australian Government."\parencite{iggulden2017}
\end{tcolorbox}

This crucial concession reveals a very different level of confidence in the pumpjet being suitable for the Australian diesel-electric submarine.  It also gives an important glimpse into the processes that selected the winning bidder, and key technology components, which reveals a clear contradictions between what would have been previously foregone conclusions.  By revealing that the pumpjet has not yet been firmly decided upon, it is directly implied that France's exclusive offer to transfer the pumpjet technology could not have been the real distinguishing feature that elevated that bid above the others.  Simultaneously, the statement that technology choice is a product of the operational requirement makes it clearly implies that the operational requirement has not yet been determined with sufficient precision to make the suitable technology.  As a ship-builder with ample experience in both technologies, uniquely positioned as having fitted both pumpjets and propellers to both diesel and nuclear submarines, it is remarkable that the performance information transferred prior or during the CEP was not sufficient to allow a decision  on the technology to be made.  Moreover, it also reveals that this was not only the case at the conclusion of the CEP, and remained so well over a year after.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{25 Oct 2017 - Herve Guillou, CEO Naval Group}]
\textit{In an interview that was published in a post\parencite{guillou2017} on the ASPI Strategist the global CEO of Naval group responds to a question from Brendan Nicholson asking whether the Australian submarine will be fitted with a pump jet, following Jean-Michel Billig's comments. Emphasis added.}
\\
\\
``Naval Group is the manufacturer of both conventional propeller-driven submarines and larger nuclear submarines with the propulsor. The propulsor's inclusion means the future submarine will move more quietly than submarines with obsolete propeller technology. \textbf{In a contest between two otherwise identical submarines, the one with the pump jet always has the tactical advantage. The propulsor was an integral part of our bid. The Department of Defence says it intends to use the pump jet propulsion system.}"
\end{tcolorbox}

The verbatim quotation of the carefully worded comparison between ``otherwise identical" submarines by the global CEO more than a year after they first appeared on the Australian DCNS website demonstrate a high degree of preparation, and continued careful word selection on the topic of the pumpjet.  This statement also further reveals nuances about the CEP selection process, simultaneously declaring that the propulsor (i.e. pumpjet) was ``integral" to their bid, whilst also deferring to what the Department of Defence ``intends" do use. This further exposes a clear tension between how 'integral' the pumpjet was to the bid, and Defence's freedom to choose otherwise.  It also exposes a dramatic divergence between the level of emphasis that is given to the pumpjet as an important key feature of the proposal in public promotion (as highlighted in \cite{lane2017}) and a Defence decision which seemed to choose only to have the option of including the pumpjet, with the real decision to be deferred.

@@

This is alarming for two reasons.  The first is that it contradicts subsequent explanations regarding pumpjet efficiency being incorrectly assessed by outsiders because they failed to account for how how the hull would be `tuned' to the pumpjet and the pumpjet `tuned' to the hull.  If the hull diameter is already fixed and known as being identical to that of the nuclear Barracuda, then no adjustment ought to be necessary to a good working pumpjet design that is used to propel this platform on the hull's account.  The performance of the pumpjet on the relevant hull would be well understood by the manufacturer offering it, and presumably available to the Australian government who were looking to acquire it.  Further discussion of the hull-propulsor interaction is given in Section \ref{hullpropulsor}.

The second is that is reveals a stunning naivety in terms of basic understanding of the hierarchy of engineering challenges.  It may be true that there are very many complex and challenging engineering calculations involved in building a submarine. However, it is certainly not true that scaling up a pressure hull to a new diameter to meet the same fixed static load is among the most difficult, risky or complex.  Whilst extending the length of a submarine is undoubtedly even easier, adjusting the diameter still to meet a structural integrity requirement to meet a known static pressure (driven by the dive depth) is a relatively simple calculation in comparison to the complexities of many other calculations involved in submarine design. For instance, re-designing the duct intake of blade arrangements of a pumpjet to maximise efficiency at a different speed whilst simultaneously attempting to avoid cavitation would be vastly more complex and uncertain given the unstable difficult-to-model dynamics of turbulence and cavitation involved.  In contrast, scaling up a hull-diameter was quite simply and successfully accomplished when the Collins Class submarine was developed by scaling up both the length and diameter of it's smaller Swedish predecessor.  In 2006, DSTO commissioned a report which was dedicated to proposition of a different hull diameter being used for submarine used to replace Collins \parencite{joubert2006} in order to achieve lower hydrodynamic drag.  If such a development path was actually thought to impose  such significant challenges on the engineering and certification process as RADM Sammut's remarks imply, there is no way that in 2006 when the replacement was presumed to be undertaken for a smaller number of new submarines which would enter service in 2026 (as opposed to mid-2030s as currently planned) that such an undertaking would have been seriously considered.

That this ranks as a significant consideration in the mind of the chief of this program should be the cause of considerable alarm to those taking an interest the project, as it would appear that he is quite alone in this belief that this really is a significant or risky hurdle in a project of this scale.  Were it known to the competitors that their ability to achieve this simple engineering requirement was in question, one might assume they would be highly offended, since any contender that was not considered competent to design a pressure hull of a new diameter would presumably never have been invited to participate.  RADM Sammut should be called upon to clarify what he meant by ``we needed to make sure that the certifying process would match the design philosophy" because if indeed he meant that Defence had already decided that their own certifying process already precluded scaled-up diameter hulls, this represents a very bold departure from what might be considered normal risk assessment processes for major engineering projects, including those used in designing and certifying the Collins.  It would also have essentially precluded both the Germans and Japanese from the outset.  Furthermore, if the advantage was considered so substantial, he should be called upon to substantiate how significant increased difficulty in certification would have been, with respect to cost and schedule, noting that the French proposal had a larger cost and longer budget than either of the other two contenders.

Among RADM Sammut's additional claims regarding the reason for choosing DCNS/Naval Group, include remarks regarding the establishment of 'sovereign capability' and 'data transfer' to Australia.  One would assume that this meant that the French had demonstrated a willingness to supply complete and timely information on key technologies, and that RADM Sammut's existing experience with the French supplying such information on the pumpjet (as Sean Costello had claimed had taken place during and prior to the conclusion of the CEP) supported this view.

@@

This comment confirms that the official policy is still to use the pumpjet, implying that the decision has indeed been made already.  This is at odds with the statement made only a few weeks earlier by Jean-Michel Billig.  This is also the first notable reference to the inter-relationship between the hull and the pumpjet, implying that the pumpjet will indeed be modified in order to suit the hull, and that the hull will be modified to suit the pumpjet.  This contradicts the notion that the same pumpjet as France is fitting to the Suffren (i.e. nuclear barracuda) will simply be included 'as is' on the Australian non-nuclear 'Shortfin' version.  Given that the length and diameter of the hull were announced as being already decided and practically identical to the nuclear version, it raises the very serious question as to whether/why the changed shape of the hull would necessitate modification of the complex pumpjet.  As pointed out in the my testimony to the Senate \ref{SENmorrison}, the emphasis of this relationship, above the vastly different operational speed profile of the nuclear and non-nuclear submarines which would affect the propulsion system profoundly, raises serious questions about the origin of the comment.

@@

This answer to a question again rests on the claim that the influence of the hull means that comparisons and studies such as those referred to in \parencite{morrison2018} can be ignored.  This is a question which will be given a substantial reply in subsequent sections, but even a quick reading superficial reading reveals alarming levels of over-reach in the answer.  (For instance, a type of ducted propeller known as a decelerating ducted propeller actually \textit{is} a pumpjet, by definition. An entire chapter (5) of the paper to which Defence is supposedly responding \parencite{morrison2018} discusses this at some length.)

@@

Bruce: I'm not sure about using these fat knobbly tires on my road-bike for the cycle tour... No-one else is using such tyres on a road-bike, I think they'd be slow and inefficient.

>>>>

However, over a number of years now the pumpjet choice has attracted sustained attention and criticism, mostly around its efficiency at low speeds, where a diesel-electric submarine would necessarily spend a great deal of its operating time.  I was commissioned by Gary Johnston of Submarines for Australia to investigate this topic in detail and produce a report\parencite{morrison2018}, which concluded that a very substantial reduction in dived range and endurance (as well as overall range) would necessarily result from the pumpjet choice.  Furthermore, that paper challenged the assumption that the pumpjet would be acoustically superior at low speeds, where open propellers would have no cavitation, and cause much less turbulence than the equivalent jet.

That paper, and a subsequent blog post\parencite{morrison2018a} on the topic which made the paper publicly available, attracted the interest of the Senate Economics Reference Committee Inquiry into the Future of Naval Shipbuilding in Australia.  Senator Chris Ketter asked written questions on notice to the Department of Defence requesting responses to the paper specifically, and related comments from public officials, with a return date of 3rd April 2018.  I was also invited to provide evidence to the committee in a public hearing on the 7th of June 2018.  In that public hearing[\ref{SENmorrison}] I stood by the conclusion of my substantive paper, which was: ``between two otherwise identical submarines, the one with the pump jet will always have a lower dived endurance, a lower dived range, a worse indiscretion ratio, a lower overall endurance and a lower overall range than the one with the propeller."  In the presence of both the outgoing acting CEO of Naval Group Australia Brent Clark, and General Manager of Submarines (for Defence) Stephen Johnson, I pointed out that that statement was yet to be challenged, and that I'd be very surprised if anyone with technical competence or professional integrity sought to challenge it.

In their own evidence during the course of the day, neither of those two did choose to challenge it.  Senator Kim Carr requested that defence produce a written response to that report for the committee prior to their handing down of their final report for the Inquiry, and Defence agreed to do their best in the time available, with the head of Capability and Sustainment Group (CASG) Kim Gillis assuring the committee that ``We  will  provide  the  committee  with  the  best  possible  advice  we can  in  the  shortest  possible  time  frame  to  meet  your  requirements  so  that  you  can  be  better informed, because we do hold that expertise." Senator Carr emphasised the importance of timeliness that ``it would look pretty bloody stupid after the report's handed in''. Despite these exhortations, in the final report of the committee this request, and Defence's acceptance of it was noted, but no acknowledgement was made of whether any response was actually received, or what it contained\parencite{SERC2018final}.

In the August edition of Australian Defence Magazine a feature article titled `Future Submarine: Pumpjet propulsor technology choice still not well explained'\parencite{kerr2018} appeared, referencing both my testimony in the hearing and the paper.  Comment for the article was sought from CASG, the Royal Australian Navy (RAN) and Naval Group, and none was forthcoming.  However, some substantial comments were provided by Dr David Kershaw, the chief of DSTG's Maritime Division, who was also Science and Technology Advisor for a time on SEA 1000, the Future Submarine Project.

@@

In early September, the Department of Defence gave a six-page written response to my paper, a full three months after the initial request. The two substantive arguments already identified were repeated, but the weight of their argument actually came to rest on another assumption.  This assumption was mentioned already by Dr Kershaw in the ADM article, but it was so completely at odds with the most basic physics that I was inclined to dismiss it as perhaps a clumsily worded statement made in an interview compounded by a little journalistic mis-quotation, rather than a substantive argument.   The re-appearance of this argument, which was spelled out and repeated multiple times in a written document that was prepared by Defence over the course of three months (well after any nominal deadline for timeliness had elapsed) dispelled all illusions that the argument was included by accident.  As such, I have also addressed it here as one of three substantial arguments made against my earlier conclusions regarding pumpjet efficiency.  This third argument coming from a document so cafeully and deliberately prepared has had a material impact on the strength of the conclusions of this paper, which seeks to consider all three lines of reasoning in sufficient detail to suitably discuss the circumstances in which such arguments might be made.

@@

\newpage
\section{A brief chronology of comment}

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{28 May 2012 - Senate Estimates}]
\textit{Senator Johnston directly challenges Rear Admiral Justin Moffit (then director of SEA 1000) about the pumpjet proposal.  Full exchange in \ref{SENMoffit}.}
\\
\\
\textbf{Senator JOHNSTON:} I am told that there are no pump jets on any conventional class submarines in the world.
\\
\textbf{Rear Adm. Moffitt:} That is quite right. It might be a very easy task for us to write it off. I do not know, because I have not done-
\\
\textbf{Senator JOHNSTON:} The French and the Russians have tried. No-one has used them. They are not being used on anything other than nuclear powered submarines-particularly the Americans.
\\
\textbf{Rear Adm. Moffitt:} \textbf{If it does not look like it makes sense, I will rule it out. But I cannot give you the answer to that yet because we have not done the work.}
\\
\textbf{Senator JOHNSTON:} Who had the idea to look at pump jets? Where did that come from? Did it come from capability group saying that we had better had a look at that?
\\
\textbf{Rear Adm. Moffitt:} I do not know the answer to that. I doubt that there is a clear answer.

@@

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{10 Feb 2016 - Sean Costello, CEO DCNS Australia}]
\textit{In a blog post\parencite{costello2016} for the ASPI Strategist prior to the announcement in April, Sean Costello describes the significance of the pump jet technology to their bid.}
\\
\\
``As a complete submarine power, we understand conventional propulsion, which is why we also understand that propulsion is but one part of the submarine puzzle. In designing to the Australian requirement, it should come as no surprise then that \textbf{the conventionally powered Shortfin Barracuda Block 1A is only 5\% lighter than its nuclear cousin.}"
\\
\\
``There's no better example of what I'm talking about than the \textbf{offer from France to transfer to Australia sovereign control and use of pump-jet propulsion technology for the Shortfin Barracuda}-technology resident only in France, the UK and the USA. Technology born from the French SSBN program a generation ago. \textbf{The stealth and hydrodynamic performances of pump jet propulsion are of course classified and in Australia known only to DCNS and the Australian Government."}

@@

The significance of the pumpjet in Frances bid was clearly acknowledged well before the public announcement of DCNS's success.  It also provides a clear indication that the size of the submarine already appeared to be quite well-known, and was only intended to be a very minor modification from the nuclear Barracuda.  For such a small variation in displacement, it might be very reasonably expected that the length of the submarine would vary (the easiest modification), but the diameter would not need to.  These comments also display quite clearly that the pumpjet technology in France's possession is recognised as common with technology in the US and the UK.  This rules out the likelihood that some special or unique pumpjet technology, better suited just to diesel-electric submarines, had been pioneered by France alone.  This comment also directly implies that information regarding the performance of the pumpjet, both its acoustic performance and propulsive performance, had already been given to the Australian  government, \textit{prior} to the conclusion of the Competitive Evaluation Process (CEP).

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{8 Apr 2016 - Sean Costello, CEO DCNS Australia, and Gerard Autret, Chief Naval Architect of the Shortfin Barracuda}]
\textit{In a blog post\parencite{autret2016} for the ASPI Strategist prior to the announcement in April, Sean Costello and Gerard Autret describe a number of the major features and design approaches for the Shortfin Barracuda, including relating the choice of the pumpjet closely to an acoustic requirement to close with another submarine at higher speed.}
\\
\\
``The challenge for any attack submarine is to maintain a nearly silent acoustic signature at a speed necessary to manoeuvre within weapons range of the target. The nuclear Barracuda is designed to \textbf{reduce radiated noise when operating at a speed sufficient in order to close a threat submarine undetected. Of particular importance is the pump-jet propulsor}, which combines a rotor and stator within a duct to significantly reduce the level of radiated noise through the effects of wake harmonisation and avoidance of cavitation."
\end{tcolorbox}

This comment provides clear context for the importance of the pump-jet.  Consistent with the physics outlined in my earlier paper \parencite{morrison2018} the acoustic advantage is only emphasised by DCNS as being significant at higher speeds.  Whether it is viable for a diesel-electric submarine (as opposed to the nuclear one mentioned here) to maintain such speeds to close with many types of targets is left unclear, or at least unmentioned.  As later discussed in my evidence to the Senate[\ref{SENmorrison}], this is consistent with DCNS being well aware with the advantages and disadvantages of the pumpjet, whilst not making verifiably false claims.

@@

The official explanation of the designer and builder's website makes a carefully worded explanation of the superiority of the pumpjet, by limiting the comparison to a ``tactical" encounter with an ``otherwise identical" submarine.  As discussed in[\ref{SENmorrison}], this is the only \textit{possible} target that an elevated speed - such as could be achieved by a diesel-electric submarine - might assist in closing with.  The submarine would be at a substantial `strategic' disadvantage in such an encounter, since its lower dived endurance, dived range, and poorer indiscretion ratio dramatically increases the probability that the pumpjet propelled submarine would wind up being the hunted, rather than the hunter.  More significantly, the speeds achieved would not allow the submarine to close with any other target from behind such as almost all surface ships, or nuclear submarines.  This statement exhibits extreme care in the wording in order to be plausibly defensible, and is arguably misleading.

Read literally, the comparison with `obsolete' propeller technology implies that the pumpjet is only quieter than obsolete propellers, rather than so much more quiet that all propellers that propellers in general are obsolete. This is consistent with my earlier findings. This is also consistent with there being care taken not to make any false claims, whilst still giving a very confident first-impression.

@@

\textbf{Rear Adm. Sammut:}  DCNS has proposed to us a conventionally-powered submarine that includes-I think you are trying to get to the form of propeller we are talking about here, Senator.

@@

\textbf{Senator CONROY:}  But what I have read is that the propulsion system is, if you like, designed and attuned to perform against the nuclear-powered aspect, and that they swap the powering out and keep the same propulsion system, but they have not adapted it in any specific way. I am inviting you to comment. The two parts work beautifully together, but the conventional power that we plan to use does not create as successful a signature at the lower speeds-so above 20 knots versus below 20 knots. Would you like to comment?

@@

\subsection{Naval Group at PACIFIC 2017 and in ASPI Strategist}
\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{4 October 2017 - Jean-Michel Billig, Executive Director Australian submarine Program, Naval Group}]
\textit{In comments to journalists at the PACIFIC 2017 trade show, the Naval Group executive in charge of the Australian program acknowledges that conventional propellers may be used on the Shortfin Barracuda \parencite{sbs2017}.  Subsequent reports quote Naval Group shifting responsibility for the propulsion choice to the Australian Government.}
\\
\\
(Interview transcribed from Reference \parencite{lane2017}.)

@@

\textbf{Jean-Michel Billig:} Yes, yes, we could, we could, we could also end up with something different
\\
\\
(End transcription from interview)
\\
\\
``However decisions on forms of propulsion are ultimately a matter for the Australian Government."\parencite{iggulden2017}
\end{tcolorbox}

This crucial concession reveals a very different level of confidence in the pumpjet being suitable for the Australian diesel-electric submarine.  It also gives an important glimpse into the processes that selected the winning bidder, and key technology components, which reveals a clear contradiction between what would have been previously foregone conclusions.  By revealing that the pumpjet has not yet been firmly decided upon, it is directly implied that France's exclusive offer to transfer the pumpjet technology could not have been the real distinguishing feature that elevated that bid above the others.  Simultaneously, the statement that technology choice is a product of the operational requirement makes it clearly implies that the operational requirement has not yet been determined with sufficient precision to make the suitable technology choice clear.  As a ship-builder with ample experience in both technologies, uniquely positioned as having fitted both pumpjets and propellers to both diesel and nuclear submarines, it is remarkable that the performance information transferred prior or during the CEP was not sufficient to allow a decision  on the technology to be made.  Moreover, it also reveals that this was not only the case at the conclusion of the CEP, and remained so well over a year after.

\begin{tcolorbox}[width=0.9\textwidth, center, colback=blue!5!white,colframe=blue!75!black, title= \textbf{25 Oct 2017 - Herve Guillou, CEO Naval Group}]
\textit{In an interview that was published in a post\parencite{guillou2017} on the ASPI Strategist the global CEO of Naval Group responds to a question from Brendan Nicholson asking whether the Australian submarine will be fitted with a pump jet, following Jean-Michel Billig's comments. Emphasis added.}
\\
\\
``Naval Group is the manufacturer of both conventional propeller-driven submarines and larger nuclear submarines with the propulsor. The propulsor's inclusion means the future submarine will move more quietly than submarines with obsolete propeller technology. \textbf{In a contest between two otherwise identical submarines, the one with the pump jet always has the tactical advantage. The propulsor was an integral part of our bid. The Department of Defence says it intends to use the pump jet propulsion system.}"
\end{tcolorbox}

The verbatim quotation of the carefully worded comparison between ``otherwise identical" submarines by the global CEO more than a year after they first appeared on the Australian DCNS website demonstrate a high degree of preparation, and continued careful word selection on the topic of the pumpjet.  This statement also further reveals nuances about the CEP selection process, simultaneously declaring that the propulsor (i.e. pumpjet) was ``integral" to their bid, whilst also deferring to what the Department of Defence ``intends" do use. This further exposes a clear tension between how `integral' the pumpjet was to the bid, and Defence's freedom to choose otherwise.  It also exposes a dramatic divergence between the level of emphasis that is given to the pumpjet as an important key feature of the proposal in public promotion (as highlighted in \cite{lane2017}) and a Defence decision which seemed to choose only to have the option of including the pumpjet, with the real decision to be deferred.

@@

This is alarming for two reasons.  The first is that it contradicts subsequent explanations regarding pumpjet efficiency being incorrectly assessed by outsiders because they failed to account for how how the hull would be `tuned' to the pumpjet and the pumpjet `tuned' to the hull.  If the hull diameter is already fixed and known as being identical to that of the nuclear Barracuda, then no adjustment ought to be necessary to a good working pumpjet design that is used to propel such a similar platform - at least not on the hull's account.  The performance of the pumpjet on the relevant hull would be well understood by the manufacturer offering it, and presumably available to the Australian government who were looking to acquire it.  Further discussion of the hull-propulsor interaction is given in Section \ref{hullpropulsor}.

The second is that is reveals a stunning naivety in terms of basic understanding of the hierarchy of engineering challenges.  It may be true that there are very many complex and challenging engineering calculations involved in building a submarine. However, it is certainly not true that scaling up a pressure hull to a new diameter to meet the same fixed static load is among the most difficult, risky or complex.  Whilst extending the length of a submarine is undoubtedly even easier, adjusting the diameter to meet a known static pressure (driven by the dive depth) is a relatively simple calculation in comparison to the complexities of many other calculations involved in submarine design. For instance, re-designing the duct intake of blade arrangements of a pumpjet to maximise efficiency at a different speed whilst simultaneously attempting to avoid cavitation would be vastly more complex and uncertain given the famously unstable, difficult-to-model dynamics of turbulence and cavitation involved.  In contrast, scaling up a hull-diameter was quite simply and successfully accomplished when the Collins Class submarine was developed by scaling up both the length and diameter of it's smaller Swedish predecessor.  In 2006, DSTO commissioned a report which was dedicated to the proposition of a different hull diameter being used for submarine used to replace Collins \parencite{joubert2006} in order to achieve lower hydrodynamic drag.  If such a development path was actually thought to impose  such significant challenges on the engineering and certification process as RADM Sammut's remarks imply, there is no way that in 2006 when the replacement was presumed to be undertaken for a smaller number of new submarines which would enter service in 2026 (as opposed to mid-2030s as currently planned) that such an undertaking would have been seriously considered.

That this ranks as a significant consideration in the mind of the chief of this program should be the cause of considerable alarm to those taking an interest the project, as it would appear that he is quite alone in this belief that this really is a significant or risky hurdle in a project of this scale.  Were it known to the competitors that their ability to achieve this simple engineering requirement was in question, one might assume they would be highly offended, since any contender that was not considered competent to design a pressure hull of a new diameter would presumably never have been invited to participate.  RADM Sammut should be called upon to clarify what he meant by ``we needed to make sure that the certifying process would match the design philosophy" because if indeed he meant that Defence had already decided that their own certifying process already precluded scaled-up diameter hulls, this represents a very bold departure from what might be considered normal risk assessment processes for major engineering projects, including those used in designing and certifying the Collins.  It would also have essentially precluded both the Germans and Japanese from the outset.  Furthermore, if the advantage was considered so substantial, he should be called upon to substantiate how significant the increased difficulty in certification would have been, with respect to cost and schedule, noting that the French proposal had a larger cost and schedule than other contenders.

Among RADM Sammut's additional claims regarding the reason for choosing DCNS/Naval Group, include remarks regarding the establishment of `sovereign capability' and `data transfer' to Australia.  One would assume that this meant that the French had demonstrated a willingness to supply complete and timely information on key technologies, and that RADM Sammut's existing experience with the French supplying such information on the pumpjet (as Sean Costello had claimed had taken place during and prior to the conclusion of the CEP) supported this view.

@@

This comment confirms that the official policy is still to use the pumpjet, implying that the decision has indeed been made already.  This is at odds with the statement made only a few weeks earlier by Jean-Michel Billig.  This is also the first notable reference to the inter-relationship between the hull and the pumpjet, implying that the pumpjet will indeed be modified in order to suit the hull, and that the hull will be modified to suit the pumpjet.  This contradicts the notion that the same pumpjet as France is fitting to the Suffren (i.e. nuclear Barracuda) will simply be included `as is' on the Australian non-nuclear `Shortfin' version.  Given that the length and diameter of the hull were announced as being already decided and practically identical to the nuclear version, it raises the very serious question as to whether/why the changed shape of the hull would necessitate modification of the complex pumpjet.  As pointed out in the my testimony to the Senate[\ref{SENmorrison}], the emphasis of this relationship, above the vastly different operational speed profile of the nuclear and non-nuclear submarines which would affect the propulsion system profoundly, raises serious questions about the origin of the comment.

@@

This answer to a question again rests on the claim that the influence of the hull means that comparisons and studies such as those referred to in Reference \parencite{morrison2018} can be ignored.  This is a question which will be given a substantial reply in subsequent sections, but even a quick superficial reading reveals alarming levels of over-reach in the answer.  (For instance, a type of ducted propeller known as a decelerating ducted propeller actually \textit{is} a pumpjet, by definition. An entire chapter (5) of the paper to which Defence is supposedly responding \parencite{morrison2018} discusses this at some length.)

@@

\textbf{Bruce:} I'm not sure about using these fat knobbly tires on my road-bike for the cycle tour... No-one else is using such tyres on a road-bike, I think they'd be slow and inefficient.
